The Jester joke recommender system

### 2020-08-02 inst/f1points.Rmd
better writing and coloured lines

The Bradley-Terry model \citep{bradley1952} assigns non-negative
strengths $p_1,\ldots, p_n$ to each of $n$ competitors in such a way
that the probability of $i$ beating $j\neq i$ in pairwise competition
is $\frac{p_i}{p_i+p_j}$; it is conventional to normalize so that
$\sum p_i=1$.  Further, we use a generalization due to
\citet{luce1959}, in which the probability of competitor $i$ winning
in a field of $j=1,\ldots, n$ is $\frac{p_i}{p_1+\cdots +p_n}$.
Noting that there is information in the whole of the finishing order,
not just the first across the line, we can follow \cite{plackett1975}
and consider the runner-up to be the winner among the remaining
competitors; and so on down the finishing order.  Without loss of
generality, if the order of finishing were $1,2,3,4,5$, then a suitable
\citeauthor{plackett1975} likelihood function would be

@@

```{r, label=calcpoints_real, echo=FALSE, cache=TRUE}
p <- formula1_points_systems()[[1]]   # real points system
dp2016 <- ordertable2points(as.ordertable(F1_table_2016[seq_len(top),]),p)
dp2016 <- sort(dp2016,decreasing=TRUE)
dp2016[] <- seq_along(dp2016)
m2016 <- maxp(H2016)
m2016 <- sort(m2016,decreasing=TRUE)
m2016[] <- seq_along(m2016)

@@

```{r,label=likevspointszipf,cache=TRUE,echo=FALSE,fig.cap="As for figure \ref{orderingbypointsandlikelihood} \\label{orderingbypointsandlikelihood_zipf} but points ranking calculated by Zipf's law.  Note the generally poorer agreement between the two ranking systems ALSO R PLOTTING SUCKS HERE AND I WILL HAVE TO CHANGE THE PDF BY HAND WITH INKSCAPE TO MAKE IT LOOK GOOD"}
par(mfrow=c(2,2))
par(mar = c(3,3,3,3))

@@

matplot(jj[wanted,],type="n",xlab='points system',ylab='mean correct places',axes=FALSE)
axis(1,at=1:6,labels=names(formula1_points_systems(top)[wanted]))
axis(2,at=1:4)
points(s2016[wanted,1],pch=1,type='b')
points(s2017[wanted,1],pch=2,type='b')
points(s2018[wanted,1],pch=3,type='b')
points(s2019[wanted,1],pch=4,type='b')
#legend("topright",lty=0,pch=1:4,legend=c("2016","2017","2018","2019"))
par(xpd=TRUE)
text(x=6,y=s2016[wanted[6],1],pos=4,"2016")
text(x=6,y=s2017[wanted[6],1],pos=4,"2017")
text(x=6,y=s2018[wanted[6],1],pos=4,"2018")
text(x=6,y=s2019[wanted[6],1],pos=4,"2019")
```

@@

matplot(jj[wanted,],type="n",xlab='points system',ylab='probability of winner correct',axes=FALSE)
axis(1,at=1:6,labels=names(formula1_points_systems(top)[wanted]))
axis(2)
points(s2016[wanted,2],pch=1,type='b')
points(s2017[wanted,2],pch=2,type='b')
points(s2018[wanted,2],pch=3,type='b')
points(s2019[wanted,2],pch=4,type='b')
#legend(x=3.4,y=0.67,lty=0,pch=1:4,legend=c("2016","2017","2018","2019"))
par(xpd=TRUE)
text(x=6,y=s2016[wanted[6],2],pos=4,"2016")
text(x=6,y=s2017[wanted[6],2],pos=4,"2017")
text(x=6,y=s2018[wanted[6],2],pos=4,"2018")
text(x=6,y=s2019[wanted[6],2],pos=4,"2019")
```

@@

matplot(jj[wanted,],type="n",xlab='points system',ylab='probability of total rank order correct',axes=FALSE)
axis(1,at=1:6,labels=names(formula1_points_systems(top)[wanted]))
axis(2)
points(s2016[wanted,3],pch=1,type='b')
points(s2017[wanted,3],pch=2,type='b')
points(s2018[wanted,3],pch=3,type='b')
points(s2019[wanted,3],pch=4,type='b')
#legend("topleft",lty=0,pch=1:4,legend=c("2016","2017","2018","2019"))
par(xpd=TRUE)
text(x=5,y=s2016[wanted[5],3],pos=3,"2016")
text(x=5,y=s2017[wanted[5],3]-0.01,pos=1,"2017")
text(x=5,y=s2018[wanted[5],3],pos=3,"2018")
text(x=5,y=s2019[wanted[5],3],pos=3,"2019")
```

>>>>

The Bradley-Terry model \citep{bradley1952} assigns non-negative
strengths $p_1,\ldots, p_n$ to each of $n$ competitors in such a way
that the probability of $i$ beating $j\neq i$ in pairwise competition
is $\frac{p_i}{p_i+p_j}$; it is conventional to normalize so that
$\sum p_i=1$.  Further, we use a generalization due to
\citet{luce1959}, in which the probability of competitor $i$ winning
in a field of $j=1,\ldots, n$ is $\frac{p_i}{p_1+\cdots +p_n}$.
Noting that there is information in the whole of the finishing order,
and not just the first across the line, we can follow
\cite{plackett1975} and consider the runner-up to be the winner among
the remaining competitors, and so on down the finishing order.
Without loss of generality, if the order of finishing were
$1,2,3,4,5$, then a suitable
\citeauthor{plackett1975}-\citeauthor{luce1959} likelihood function
would be

@@

```{r, label=calcpoints_real, echo=FALSE, cache=TRUE}
```{r, label=calcpoints_real, echo=FALSE, cache=TRUE}
p <- formula1_points_systems()[[1]]   # real points system
dp2016 <- ordertable2points(as.ordertable(F1_table_2016[seq_len(top),]),p)
dp2016 <- sort(dp2016,decreasing=TRUE)
dp2016[] <- seq_along(dp2016)
m2016 <- maxp(H2016)
m2016 <- sort(m2016,decreasing=TRUE)
m2016[] <- seq_along(m2016)

@@

	```{r,label=likevspointszipf,cache=TRUE,echo=FALSE,fig.cap="As for figure \ref{orderingbypointsandlikelihood} \\label{orderingbypointsandlikelihood_zipf} but points ranking calculated by Zipf's law. REF DOES NOT WORK HERE  Note the generally poorer agreement between the two ranking systems ALSO R PLOTTING SUCKS HERE AND I WILL HAVE TO CHANGE THE PDF BY HAND WITH INKSCAPE TO MAKE IT LOOK GOOD"}
par(mfrow=c(2,2))
par(mar = c(3,3,3,3))

@@

matplot(jj[wanted,],type="n",xlab='points system',ylab='mean correct places',axes=FALSE)
axis(1,at=1:6,labels=names(formula1_points_systems(top)[wanted]))
axis(2,at=1:4)
points(s2016[wanted,1],pch=1,type='b',col='red')
points(s2017[wanted,1],pch=2,type='b',col='orange')
points(s2018[wanted,1],pch=3,type='b',col='green')
points(s2019[wanted,1],pch=4,type='b',col='blue')
#legend("topright",lty=0,pch=1:4,legend=c("2016","2017","2018","2019"))
par(xpd=TRUE)
text(x=6,y=s2016[wanted[6],1],pos=4,"2016",col='red')
text(x=6,y=s2017[wanted[6],1],pos=4,"2017",col='orange')
text(x=6,y=s2018[wanted[6],1],pos=4,"2018",col='green')
text(x=6,y=s2019[wanted[6],1],pos=4,"2019",col='blue')
```

@@

matplot(jj[wanted,],type="n",xlab='points system',ylab='probability of winner correct',axes=FALSE)
axis(1,at=1:6,labels=names(formula1_points_systems(top)[wanted]))
axis(2)
points(s2016[wanted,2],pch=1,type='b',col='red')
points(s2017[wanted,2],pch=2,type='b',col='orange')
points(s2018[wanted,2],pch=3,type='b',col='green')
points(s2019[wanted,2],pch=4,type='b',col='blue')
#legend(x=3.4,y=0.67,lty=0,pch=1:4,legend=c("2016","2017","2018","2019"))
par(xpd=TRUE)
text(x=6,y=s2016[wanted[6],2],pos=4,"2016",col='red')
text(x=6,y=s2017[wanted[6],2],pos=4,"2017",col='orange')
text(x=6,y=s2018[wanted[6],2],pos=4,"2018",col='green')
text(x=6,y=s2019[wanted[6],2],pos=4,"2019",col='blue')
```

@@

matplot(jj[wanted,],type="n",xlab='points system',ylab='probability of total rank order correct',axes=FALSE)
axis(1,at=1:6,labels=names(formula1_points_systems(top)[wanted]))
axis(2)
points(s2016[wanted,3],pch=1,type='b',col='red')
points(s2017[wanted,3],pch=2,type='b',col='orange')
points(s2018[wanted,3],pch=3,type='b',col='green')
points(s2019[wanted,3],pch=4,type='b',col='blue')
#legend("topleft",lty=0,pch=1:4,legend=c("2016","2017","2018","2019"))
par(xpd=TRUE)
text(x=5,y=s2016[wanted[5],3],pos=3,"2016",col='red')
text(x=5,y=s2017[wanted[5],3],pos=1,"2017",col='orange')
text(x=5,y=s2018[wanted[5],3],pos=3,"2018",col='green')
text(x=5,y=s2019[wanted[5],3],pos=3,"2019",col='blue')
```
